Generalized Plackett-Luce Likelihoods

### 2022-08-12 inst/jss_hyper3.Rnw
point to github repo
rewording

The ideas presented above can easily be extended to arbitrarily large
numbers of competitors, although the analytical expressions tend to be
intractable and numerical methods must be used.  Given an order
statistic of the type considered above, the Mann-Whitney-Wilcoxon
test~\citep{mann1947,wilcoxon1945} assesses a null of identity of
underlying distributions.  Consider the chorioamnion dataset at
\code{wilcox.test.Rd}:

@@

A particularly attractive feature of this analysis is that the
Bradley-Terry strengths have direct operational significance: If two
competitors, say Vadlejch and Vesely, were to throw a javelin, then we
would estimate the probability that competitors, say Vadlejch would
throw further than Vesely at $\displaystyle p_{\mbox{\tiny
Vad}}/\left(p_{\mbox{\tiny Vad}} + p_{\mbox{\tiny Ves}}\right)\simeq
0.74$.  Indeed, from a training or selection perspective we might
follow \cite{hankin2017} and observe that log-contrasts
~\citep{ohagan2004} $\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny
Ves}}\right)$ appear to have approximately Gaussian likelihood
functions for observations of the type considered here.  Profile
log-likelihood curves for log-contrasts are easily produced by the
package, Figure~\ref{profliklogcont}.

>>>>

The ideas presented above can easily be extended to arbitrarily large
numbers of competitors, although the analytical expressions tend to be
intractable and numerical methods must be used.  All results and
datasets presented here are maintained under version control and
available at \url{https:github.com/RobinHankin/hyper2}.  Given an
order statistic of the type considered above, the
Mann-Whitney-Wilcoxon test~\citep{mann1947,wilcoxon1945} assesses a
null of identity of underlying distributions.  Consider the
chorioamnion dataset at \code{wilcox.test.Rd}:

@@

A particularly attractive feature of this analysis is that the
Bradley-Terry strengths have direct operational significance: If two
competitors, say Vadlejch and Vesely, were to throw a javelin, then we
would estimate the probability that Vadlejch would throw further than
Vesely at $\displaystyle p_{\mbox{\tiny Vad}}/\left(p_{\mbox{\tiny
Vad}} + p_{\mbox{\tiny Ves}}\right)\simeq 0.74$.  Indeed, from a
training or selection perspective we might follow \cite{hankin2017}
and observe that log-contrasts~\citep{ohagan2004}
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$ appear
to have approximately Gaussian likelihood functions for observations
of the type considered here.  Profile log-likelihood curves for
log-contrasts are easily produced by the package,
Figure~\ref{profliklogcont}.
